---
title: "RNN on Weather Forecasting"
author: "Rajendra"
output: html_document
---


Libraries
```{r,message=FALSE}
library(tibble)
library(readr)
library(keras)
```

Importing Dataset
```{r,message=FALSE}
dataset <- read_csv("jena_climate_2009_2016.csv")
```

Data Preperation
Dimensions of the data
```{r}
dim(dataset)
```

sampling the dataset to fewer samples (20000 samples) and excluding 1st coloumn which have data and time for further analysis.

```{r,message=FALSE}
data <- dataset[1:20000,]
data <- data.matrix(data[,-1])
```


Data Preprocessing
Partitioning the data into **train,test & validation splits**
```{r}
train_data <- data[1:15000,]
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)

# Scaling Data
data <- scale(data, center = mean, scale = std)
```


**Generator**

```{r}
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 128, step = 6) {
  if (is.null(max_index))
    max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size, max_index))
      i <<- i + length(rows)
    }
    samples <- array(0, dim = c(length(rows),
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]],
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,2]
    }
    list(samples, targets)
  }
}
```



```{r}
lookback <- 1440
step <- 6
delay <- 144
batch_size <- 128
```


```{r}
train_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 15000,
  shuffle = TRUE,
  step = step,
  batch_size = batch_size
)
val_gen = generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 15001,
  max_index = 17000,
  step = step,
  batch_size = batch_size
)
test_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 17001,
  max_index = 19000,
  step = step,
  batch_size = batch_size
)

# Caluclating the valuations steps
val_steps <- (17000 - 15001 - lookback) / batch_size
# Caluclating the test steps
test_steps <- (19000 - 17001 - lookback) / batch_size
```


 Model Construction
Let's evaluate this approach, using the mean absolute error(MAE) metric
```{r}
evaluate_naive_method <- function() {
  batch_maes <- c()
  for (step in 1:val_steps) {
    c(samples, targets) %<-% val_gen()
    preds <- samples[,dim(samples)[[2]],2]
    mae <- mean(abs(preds - targets))
    batch_maes <- c(batch_maes, mae)
  }
  print(mean(batch_maes))
}
evaluate_naive_method()
```



Densely Connected Model
Training and evaluating a densely connected model
```{r}
model <- keras_model_sequential() %>%
  layer_flatten(input_shape = c(lookback / step, dim(data)[-1])) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```
```{r,echo=FALSE,fig.align="Center"}
plot(history)
```

from the above plot we see that validation losses are similar to base model


Recuurent Network model with stacked *layer_gru*


```{r}
model <- keras_model_sequential() %>% 
  layer_gru(units = 64, 
            dropout = 0.1, 
            recurrent_dropout = 0.5,
            return_sequences = TRUE,
            input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_gru(units = 128, activation = "relu",
            dropout = 0.1,
            recurrent_dropout = 0.5) %>% 
  layer_dense(units = 1)
model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)
history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 50,
  epochs = 5,
  validation_data = val_gen,
  validation_steps = val_steps
)
```
```{r,echo=FALSE,fig.align="Center"}
plot(history)
```


overfitting is stopped during the first 5 epochs.


Evaluating on test set
```{r}
result <- model %>% fit_generator(
  test_gen,
  steps_per_epoch = 50,
  epochs = 5,
  validation_steps = test_steps
)
```
```{r,echo=FALSE}
plot(result)
```

By the above plot we can see that stacked model is performing well on the test data which is less than base model 